---
title: "Practical Time Series Analysis"
subtitle: "Thistleton and Sadigov"
format: html
keywords: time series analysis, forecasting, R, linear regression, normality, residuals, t-test, coursera
#draft: true
---

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(faraway)

```

I took this course on Coursera in primarily to get a more solid foundation in 
time series analysis and forecasting before taking the bayesian time series
course. I have some experience with time series analysis from reading a few
books and papers, but I wanted to get a more solid foundation in the subject.

Most of the material in this course is introductory and I did not bother 
with very in depth explanations or notes, often listing just what is covered.


reading for week 1:

- [Download-and-Install-for-Windows](w1_Introduction-to-R---Download-and-Install-for-Windows.pdf)
- [Getting-Started-with-Packages](w1_Introduction-to-R---Getting-Started-with-Packages-in-R.pdf)
- [Reviewing-Basic-Statistics](w1_Introduction-to-R---Reviewing-Basic-Statistics.pdf)
- [Measuring Linear Association](w1_Measuring-Linear-Association-with-the-Correlation-Function.pdf)

# Module 1

- We begin with the usual pleasantries, and
- R software installation
- How to load a library 
- How to load a dataset from a library
- Getting some info on the dataset using e.g. `help(c02)` for the Mauna Loa CO2 dataset.
- How to get a five point summary using `summary` function
- How to visualize a dataset with a **histogram** in R with the `hist` function
  - Customize title, x-label, y-label, color, and breaks
  - How to add a density plot to a histogram
- How to visualize a bivariate dataset with a **scatter plot** in R using `plot` function
  - Customize title, x-label, y-label, 
  - How to add a regression line to a scatter plot with the `abline` function
- How to fit a liner regression model to a dataset in R using `lm` function
  - How we interpret the output of the `lm` function
  - How to get the coefficients of the model using `coef` function
  - Understanding the error term in the regression model
  - Talks about some logical assumptions of the linear regression model ^[these are 
    not logical simply some of the mathematical requirements made when estimating 
    least squares.]
    - Normally distributed errors
    - same variance of errors (homoscedasticity)
    - independent errors (no autocorrelation)
  - using `()` around an expression to coerce the output to the console

```{r}
require(graphics)
summary(co2)
co2lm = lm(co2 ~ time(co2))

# plot data and fit
plot(co2, main='CO2 data', xlab='Time', ylab='CO2')
abline(co2lm, col='red')
co2lm.residuals = residuals(co2lm)

# check the residuals
plot(co2lm.residuals~time(co2), main='Residuals', xlab='Time', ylab='Residuals')

# check the normality of the residuals
qqnorm(residuals(co2lm))
qqline(residuals(co2lm))
```


```{r}

library(faraway)
plot(punting$Distance~ punting$Hang); 
abline( lm(punting$Distance~ punting$Hang) )
m=lm(punting$Distance~ punting$Hang) 
qqnorm(resid(m)) 
qqline(resid(m))
```

- using `t.test`

```{r}
attach(sleep)
plot(extra~group,data=sleep,main="Extra Sleep in Gossett Data by Group",xlab="Group",ylab="Extra Sleep")
extra_1=extra[group==1]
extra_2=extra[group==2]
t.test(extra_1,extra_2,paired=TRUE,alternative="two.sided")

```
the confidence interval does not contain zero, so we can reject the null hypothesis that the means are equal, and conclude that the means are different.

## Visualization homework

Which of the following has the right R command and the right histogram for the dataset (called quiz_data) from Part 1 (provided below) with a title 'Histogram', x-label 'Quiz data' and 10 break points?

37,86,79,95,61,93,19,98,121,26,39,11,26,75,29,130,42,8



```{r}
quiz_data=c(37, 86, 79, 95, 61, 93, 19, 98, 121, 26, 39, 11, 26, 75, 29,130, 42, 8)
hist(quiz_data,breaks=10,main='Histogram',xlab='Quiz data')
```

You can add options to executable code like this 

```{r}
hist(quiz_data, freq=F, breaks=10, main='Histogram', xlab='Quiz data', col='blue')
lines(density(quiz_data), col='red', lwd=5)

```

The `echo: false` option disables the printing of code (only output is displayed).
